---
title: "Data Programming with Python"
author: "Hannah Luebbering"
date: "July 04, 2022"
output: 
  html_document: 
    css: "assets/main.css"
    toc: yes
    toc_float: yes
knit: (function(inputFile, encoding) {rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(kableExtra)
library(knitr)

source("data.R")
```


## Introduction

The goal of this tutorial is to gain more experience writing and running Python code and programs and practice printing results to match a specification.

### Data Types

For each expression, we write the resultant value and data type as shown in the below table.

```{r}
kable(df1, col.names = c("Expression", "Resultant Value", "Data Type")) %>%
  kable_styling(full_width = F, font_size = 12, html_font = "Roboto Condensed", bootstrap_options = c("striped", "hover"))
```


### Lists + Range


For each list below, we write an equivalent call using `range()`. Similarly, for each call using range, we give the corresponding list.


```{r}
kables(list(kbl2, kbl3)) %>%
  kable_styling(full_width = F, font_size = 12, html_font = "Roboto Condensed", bootstrap_options = "none")
```







-------------------------


## Loops + If Statements

In the following section, we write Python code using variables and loops to solve simple math problems.


### Problem 1

<div class = "datasource2">
<ol><li value = "1"> Compute and print both roots in ascending order of the following quadratic equation: $3x^2-5.86x+2.5408$</li></ol>
</div>


Here we need to use the quadratic equation to solve for the roots:

$$
\begin{align}
x &= \frac{-b\pm \sqrt{b^2 - 4ac}}{2a} \\
&= \frac{5.86 \pm \sqrt{5.86^2 - 4\cdot 3 \cdot 2.5408}}{2\cdot 3}
\end{align}
$$



```python
import math

a = 3
b = -5.86
c = 2.5408
root1 = (-b - math.sqrt((b ** 2) - (4 * a * c))) / (2 * a)
root2 = (-b + math.sqrt((b ** 2) - (4 * a * c))) / (2 * a)
print("Root 1: ", root1)  # 0.6496431
print("Root 2: ", root2)  # 1.30369
```


-------------------------


### Problem 2

<div class = "datasource2">
<ol><li value = "2"> Use a for loop to print the decimal representations of the following fractions, each on their own line: $\left\{\frac{1}{2},\frac{1}{3}, \ldots , \frac{1}{10} \right\}$. </li></ol>
</div>


```python
for denom in range(2, 11, 1):
    print("1 /", denom, ":", float(1/denom))
```


-------------------------


### Problem 3 

<div class = "datasource2">
<ol><li value = "3">
Use a for loop to compute the $10$th triangular number. The $n$th triangular number is calculated by $\frac{n(n+1)}{2}$.
</li></ol>
</div>


The following solution should correctly calculate the $11$th, $12$th, $\ldots$,  triangular number by changing the variable $n$ for any number greater than $0$.


```python
n = 10
triangular = 0
for i in range(1, n + 1):
    triangular += i
print("Triangular number", n, "via loop:", triangular)
print("Triangular number", n, "via formula:", n * (n + 1) / 2)
```


-------------------------

### Problem 4

<div class = "datasource2">
<ol><li value = "4">
Use a for loop to compute $10!$, the factorial of $10$. The factorial of a number, $n$, can be calculated by $1\cdot 2 \cdot 3 \cdot \ldots \cdot n$.
</li></ol>
</div>


Similarly to problem 3, the following solution should be able to correctly calculate any factorial just by changing a variable.


```python
n = 10
factorial = 1
for i in range(1, n + 1):
    factorial = (factorial * i)
print(factorial)
```



-------------------------

### Problem 5

<div class = "datasource2">
<ol><li value = "5">
Print the first $10$ factorials in descending order $\left(10!, 9!,\ldots , 1!\right)$.
</li></ol>
</div>

The first line of your solution should assign a variable num_lines to 10, and then the rest of your solution should print the correct number of lines and factorial on its own line. Similarly to problems 3 and 4, your solution should be able to work for any number just by changing num_lines.


```python
num_lines = 10
for num in range(num_lines, 0, -1):
    factorial = 1
    for i in range(1, num + 1):
        factorial = (factorial * i)
    print(factorial)
```


-------------------------


### Problem 6

<div class = "datasource2">
<ol><li value = "6">
Use a for loop to compute the following value:
</li></ol>
</div>



Similar to problems 3 - 5, you should assign a variable to determine how many number of fractions to add. Don’t try to incorporate adding 1 into your loop as that will unnecessarily complicate your solution. Instead, add it to the result either prior or after running your for loop.

```python
fractions = 10
value = 0
for num in range(0, fractions, 1):
    factorial = 1
    for i in range(1, num + 1):
        factorial = (factorial * i)
    value += 1/factorial
print(value)  # 2.718281801
```

-------------------------


## Functions + Lists


<span class = "myhighlight">Learning objective:</span> Read, process, and group CSV data to compute descriptive statistics with and without Pandas.

- `hw2_manual.py` is the file for you to put your implementations for solving each problem without Pandas.
- `hw2_pandas.py` is the file for you to put your implementations for solving each problem with Pandas.
- `hw2_test.py` is the file for you to put your own tests. The Run button executes this program.
- `cse163_utils.py` is a helper file that has code to help you test your code.
- `pokemon_box.csv` is a large CSV file that stores information about pokemon.
- `pokemon_test.csv` is a very small CSV file that stores information about pokemon used for the example cases.
- `PandasWordbank.ipynb` is a Jupyter Notebook that reviews all of the Pandas features needed for this assessment. Feel free to edit this Jupyter Notebook to prototype ideas and explore the data.
- `pets.csv` is the tiny example dataset used only by the PandasWordbank.ipynb


-------------------------




## File IO + Dictionaries




## Sets & Sorting




-------------------------








